On the Synchronization Power of the Openflow Data Plane

### 2015-04-16 inband-sync.tex
moved two algs to appendix "to be deleted"
some updates to bundle subsection

The Openflow bundle can be seen as a ``mini-transaction'':
a bundles aggregates multiple control commands (e.g., FlowMod messages)
into one execution, so that either all of the control messages are
processed, or none of them is. In particular, if, for some reason, the
configuration request contained in one of the bundled messages cannot
be processed, all the other requests  are rejected and an error
message is sent to the controller that issued them.
This bundled processing can be, when properly configured, done in a
\emph{packet-atomic} way, so that each packet is processed according
to the switch configuration either \emph{before} or \emph{after} the
request was processed.
If multiple FlowMod messages with the CHECK\_OVERLAP flag are sent as a single bundle, the bundle command also returns a reason 
if the checks failed: the failed command in the bundle.
%:a CHECK\_OVERLAP flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}

@@

We will make use of bundling only for policy updates. 
Bundling is \emph{not required} to implement transactions.

In the remainder of this paper, we will use the notation 
$\textbf{bundle}\{\;\ldots\;\}$ to refer to represent a
sequence of commands in a bundle.
\stefan{liron: i think the following needs more details $:-)$ :} A bundle will always have the following structure:
it first contains a \emph{create bundle message} to start bundle,
followed by a sequence of messages (e.g., FlowMod) each wrapped in a \emph{add to bundle message} with the bundle ID, 
and finally \emph{commit bundle} control message for commit.
A bundle always returns a binary result: \emph{error} or \emph{success}.

@@

To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
message from the switch is received. When the switch receives the
commit request it tries to atomically process all the buffered messages
equipped with the corresponding bundle identifier and, if the
processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
message  pointing out the aborted transactional operation.

Algorithm~\ref{alg:commit} describes how to commit
the bundle. 
\begin{algorithm}[h]
    \caption{$\textit{try-commit}()$}
    \label{alg:commit}
    \begin{algorithmic}[1]
    \Require  current bundle id $bid$.
		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
    		\State  $cmd.bundle\_id \gets bid$
    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST}
    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
    		\State send $cmd$
		\Return $cmd$
    \end{algorithmic}
\end{algorithm}

%\textbf{Cookies:} Cookies allow to name and retrieve individual
%flow entries and to use masks to delete modify/delete multiple entries. 
%Our synchronization mechanism will implement controller-specific cookies.
%
%\item \textbf{Groups:} Groups of action buckets can be defined in a special group table. Each group has a unique id and any attempt to create a group with existing id or %to delete un existing group fails (and can abort a bundle in process).
%\end{enumerate}

@@

\begin{algorithm}[t]
    \caption{Policy composition without bundle}
    \label{alg:wobundle}
    \begin{algorithmic}[1]
    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
    %\Ensure installed policy is consistent with previous one
    \State $unactive \gets active + 1 \mod{2}$
    \For {$cmd \in P$}
	    \State $cmd.action\gets Reencode(cmd.action)$
	    \State $cmd.table\_id\gets T_{unactive}$
	    \State send $cmd$
    \EndFor
    \State modify the action of $T_s$ entry to goto $T_{unactive}$
    \State clear table $T_{active}$
    \State $active \gets unactive$
	\Return

\begin{algorithm}[t]
    \caption{Update with pipeline}
    \label{alg:pipeline}
    \begin{algorithmic}[1]
    \Require update requirements $U$, array $ARR$.
    \Ensure installed policy is consistent with previous one
			\Return

@@

>>>>

The Openflow bundle can be seen as a ``mini-transaction'':
a bundles aggregates multiple control commands (e.g., FlowMod messages)
into one execution, so that either all of the control messages are
performed, or none of them is. In particular, if, for some reason, the
configuration request contained in one of the bundled messages cannot
be performed, all the other requests  are rejected and an error
message is sent to the controller that issued them.
This bundled processing can be, when properly configured, done in a
\emph{packet-atomic} way, so that each packet is processed according
to the switch configuration either \emph{before} or \emph{after} the
request was processed.
If multiple FlowMod messages with the CHECK\_OVERLAP flag are sent as a single bundle, the bundle command also returns a reason 
if the checks failed: the failed command in the bundle.
%:a CHECK\_OVERLAP flag can be defined for each command in the bundle.\liron{these last two sentences are not cleaar}

@@

We will make use of bundling only for policy updates. 
Bundling is \emph{not required} to implement transactions.
\liron{what does is means?}

In the remainder of this paper, we will use the notation 
%$\textbf{bundle}\{\;\ldots\;\}\rightarrow res$ to refer to represent a
$\textbf{bundle}\{\;\ldots\;\}$ to refer to represent a
sequence of commands sent in a bundle, returning $True$ if and only if all are performed successfully.
\stefan{liron: i think the following needs more details $:-)$ :}\liron{is it ok now?} A bundle will always have the following structure:
it first contains a \emph{create bundle message} ( \textsf{OFPT\_BUNDLE\_CONTROL} message with type
\textsf{OFPBCT\_OPEN\_REQUEST}) to start bundle,
followed by a sequence of commands (e.g., FlowMod) each wrapped in a \emph{add to bundle message} ( \textsf{OFPT\_BUNDLE\_ADD\_MESSAGE}) with the bundle ID, 
and finally \emph{commit bundle} control message (\textsf{OFPT\_BUNDLE\_CONTROL} message with type \textsf{OFPBCT\_COMMIT\_REQUEST}) for commit.
A bundle always returns a binary result: \emph{error} or \emph{success}.

@@

\hide{
To commit the transaction,  the controller sends a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
\textsf{OFPBCT\_COMMIT\_REQUEST} and waits until a matching control
message from the switch is received. When the switch receives the
commit request it tries to atomically process all the buffered messages
equipped with the corresponding bundle identifier and, if the
processing is successful for all of them, returns  a \textsf{OFPT\_BUNDLE\_CONTROL} message with type
\textsf{OFPBCT\_CLOSE\_REPLY}; otherwise it sends a specific error
message  pointing out the aborted transactional operation.

Algorithm~\ref{alg:commit} describes how to commit
the bundle. 
\begin{algorithm}[h]
    \caption{$\textit{try-commit}()$}
    \label{alg:commit}
    \begin{algorithmic}[1]
    \Require  current bundle id $bid$.
		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
    		\State  $cmd.bundle\_id \gets bid$
    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST}
    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
    		\State send $cmd$
		\Return $cmd$
    \end{algorithmic}
\end{algorithm}
}
%\textbf{Cookies:} Cookies allow to name and retrieve individual
%flow entries and to use masks to delete modify/delete multiple entries. 
%Our synchronization mechanism will implement controller-specific cookies.
%
%\item \textbf{Groups:} Groups of action buckets can be defined in a special group table. Each group has a unique id and any attempt to create a group with existing id or %to delete un existing group fails (and can abort a bundle in process).
%\end{enumerate}

@@

@@

\section{To Be Deleted...}

\begin{algorithm}[h]
    \caption{Policy composition without bundle}
    \label{alg:wobundle}
    \begin{algorithmic}[1]
    \Require policy $P$ (a set of FlowMod entry creation commands), three consecutive flow tables ids (per controller) $T_s,T_0,T_1$, active table index $active\in\{0,1\}$.
    %\Ensure installed policy is consistent with previous one
    \State $unactive \gets active + 1 \mod{2}$
    \For {$cmd \in P$}
	    \State $cmd.action\gets Reencode(cmd.action)$
	    \State $cmd.table\_id\gets T_{unactive}$
	    \State send $cmd$
    \EndFor
    \State modify the action of $T_s$ entry to goto $T_{unactive}$
    \State clear table $T_{active}$
    \State $active \gets unactive$
	\Return

\begin{algorithm}[h]
    \caption{Update with pipeline}
    \label{alg:pipeline}
    \begin{algorithmic}[1]
    \Require update requirements $U$, array $ARR$.
    \Ensure installed policy is consistent with previous one
			\Return
